---
title: "01_Munging" 
author: "Jessica Stephens"
date: "10/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing Data

data is in a microsoft excel workbook
readxl

```{r load, echo=FALSE}

#check wd and file path match
setwd("C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Treatment_Aging_1yrBands_FY20")

#read xlsx - specify path, sheet name to read

df <- read_xlsx("Data/HIV2020Estimates_ART by single year_8Sept2020_norow.xlsx",
                sheet="DataList") 


```


## Check Data Import


```{r load, echo=FALSE}

glimpse(df)
View(df)

```




## Clean dataset: Isolate Age from E_Ind and drop unnecessary columns/variables

Goal: the numbers for each one-year age band of people with HIV and of people on ART
# make a new variable with the the values after "Age" and before ";"
```{r load, echo=FALSE}



df_full_seperate <- df %>% 
    separate(E_Ind, c("A","B", "C", "D", "E"), sep = "([+;])") %>%#"B" holds the age
    separate(B, c("X", "Y", "Age_num"), sep = "([ ])")%>%
    separate(A, c("X1", "PLVHIV_or_ART", "X2"), sep = "([-])")%>%
    separate("PLVHIV_or_ART", c("PLVHIV_or_ART", "Y1"), sep = "([0])") %>% #drop 0 from PLHIV vs ART
    rename(Country=E_Count, sex=D, sex2=E)  %>%#rename country variable and vars with sex
    mutate(malefemale=ifelse(is.na(sex2), 0, 1)) %>%
    mutate(sex=ifelse(malefemale==1, "MaleFemale", sex)) %>% # not working to change sex here %>%
    mutate(Country=recode(Country, "Cote dIvoire"="Cote d'Ivoire", 
                          "Lao People Democratic Republic"="Laos",
                          "United Republic of Tanzania"= "Tanzania",
                          "Viet Nam"="Vietnam"))%>%
    select(Country, Age_num, sex, Time, Value) %>% #drop X1, X2, Y, X, C,Y1, sex2 Rounded
    view() %>%
    glimpse()
  

df_full_seperate$Age_num <-
    as.numeric(as.character(unlist(df_full_seperate$Age_num)))

view(df_full_seperate)
glimpse(df_full_seperate)

```






#create FY20 columns
```
df_spec <- df_full_seperate %>% 
  mutate("FY20Q1" = NA, "FY20Q2" = NA, "FY20Q3" = NA, "FY20Q4"= NA)%>% 
  rename(FY19Q4=Value)  %>% 
  filter(Time %in% c('2019'))  %>% 
    select(-c(Time))
  view(df_spec)
  glimpse(df_spec)
```




#reshape FY20 long, breaking out qtr
```
df_pred <- df_spec %>% 
  pivot_longer(starts_with("FY20"), 
               names_to = c("fy", "qtr"), 
               names_sep = "Q",
               names_transform = list(qtr = as.integer))
               
view(df_pred)
```



#create the predicted quarterly values
```
df_pred2 <- df_pred %>% 
  group_by(Country, sex, qtr) %>% 
  mutate(sub = FY19Q4 * (.25 * qtr),  #subtracting .25 of current age value each quarter (current aka  fy19q4)
         add = lag(FY19Q4)* (.25 * qtr), #add .25 of previous age each quarter
         add = ifelse(is.na(add), 0, add)) %>% 
  ungroup() %>% 
  mutate(value = FY19Q4 + add - sub) %>% 
  select(-c(sub, add))
  
  view(df_pred2)
```




#gather then re-spread to get full time series 
```
df_pred3 <- df_pred2 %>%
  pivot_wider(names_from = c(fy, qtr),
              names_sep = "Q") 
  view(df_pred3)            
              
df_pred4 <- df_pred3 %>%
  pivot_longer(-c(Age_num, sex, Country),
               names_to = "period")
  view(df_pred4)        
  
               
```

#find max age
```

range(df_pred4$Age_num)

#0-80
```


#create age groups with cutpoints
```
library(data.table)

PEPFARagebreaks19 <- c(0,1,5,10,15,20,25,30,35,40,45,50,85)
PEPFARagelabels19 <- c("<01","01--04", "05--09","10--14","15-19","20-24","25-29","30-34",
               "35-39","40-44","45-49","50+")


df_pred5 <- setDT(df_pred4) [ , Age_cat := cut(Age_num, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)]


view(df_pred5)

glimpse(df_pred5)

names(df_pred5)



```

#create correction factor
```
df_corr <- df_pred5 %>% 
  group_by(Age_num, Country, sex) %>% 
  mutate(adj_fctr = value/lag(value)) %>% 
  ungroup()
  view(df_corr)

```

#format to merge with MSD
```
df_corr_merge<-df_corr%>%
  select(Country, period, sex, Age_num, Age_cat, adj_fctr)%>%
  rename(ageasentere=Age_num, age_group=Age_cat)%>%
  mutate_if(is.factor, as.character)%>%
  mutate(sex=ifelse(sex==" Female", "Female", sex)) %>%
  mutate(sex=ifelse(sex==" Male", "Male", sex)) %>%
  glimpse()

view(df_corr_merge)
glimpse(df_corr_merge)


write_csv(df_corr_merge,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_qtr_20201217_FY20.csv", na="")
```























#############################################
# Restructure data
##############################################

## add variables for Age Categories for ACTUAL Age

Create categories for PEPFAR age brackets
<1
1-4
5-9
10-14
15-19
20-24
25-29
30-34
35-39
40-44
45-49
50+


https://stackoverflow.com/questions/12979456/r-code-to-categorize-age-into-group-bins-breaks 


```{r load, echo=FALSE}
#find max
range(df_full_seperate$Age_num)
#0-80

library(data.table)

PEPFARagebreaks19 <- c(0,1,5,10,15,20,25,30,35,40,45,50,85)
PEPFARagelabels19 <- c("<01","01--04", "05--09","10--14","15-19","20-24","25-29","30-34",
               "35-39","40-44","45-49","50+")


df_full_PEPFARage2019 <- setDT(df_full_seperate) [ , ACTUAL_Age_cat_SemiFine := cut(Age_num, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)]


view(df_full_PEPFARage2019)

glimpse(df_full_PEPFARage2019)

names(df_full_PEPFARage2019)

```




# Figure out predicted ages 
#ex, 2017 ages if 2016 aged w/o other loss/gains


```{r load, echo=FALSE}
### annual code
# 
# df_full_PREDICTagetime <- df_full_PEPFARage2019 %>%
#  mutate(Age_predict=Age_num+1, 
#          Time_predict=Time+1)
# 
# view(df_full_PREDICTagetime)
# df_full_PREDICTagegroups <- setDT(df_full_PREDICTagetime) [ , PREDICT_Age_cat_SemiFine := cut(Age_predict, 
#                                 breaks = PEPFARagebreaks19, 
#                                 right = FALSE, 
#                                 labels = PEPFARagelabels19)]
# view(df_full_PREDICTagegroups)
# glimpse(df_full_PREDICTagegroups)



###2nd attempt at qtr aging - wide...
# 
# df_full_PREDICTagetime_qtr <- df_full_seperate %>%
#  mutate(Value_predict_qtr1=Value*.25,
#         Value_predict_qtr2=Value*.5,
#         Value_predict_qtr3=Value*.75,
#         Value_predict_qtr4=Value,
#         Age_predict=Age_num+1,
#         Time_predict=Time+1)
# view(df_full_PREDICTagetime_qtr)



### 1st attempt at mo aging

# df_full_PREDICT_expand <-
#   bind_rows(replicate(12,df_full_PEPFARage2019, simplify = FALSE),.id ="mo" ) %>%
#   mutate(Value=Value/12, 
#         qtr=ifelse(mo %in% 1:3, 1,
#             ifelse(mo %in% 4:6, 2, 
#             ifelse(mo %in% 7:9, 3, 4))))
# view(df_full_PREDICT_expand)


df_full_PREDICT_expand <-
  bind_rows(replicate(4,df_full_PEPFARage2019, simplify = FALSE),.id ="qtr" ) %>%
  mutate(Value=Value/4) 
view(df_full_PREDICT_expand)

df_full_PREDICTagetime_qtr <- df_full_PREDICT_expand %>%
 mutate(Age_predict_qtr1=
          ifelse(qtr %in% 4,Age_num+1, Age_num),
        Age_predict_qtr2=
          ifelse(qtr %in% 3:4,Age_num+1, Age_num),
        Age_predict_qtr3=
          ifelse(qtr %in% 2:4,Age_num+1, Age_num),
        Age_predict_qtr4=Age_num+1,
        Time_predict=Time+1)
view(df_full_PREDICTagetime_qtr)



df_full_PREDICTagegroups_qtr1 <- setDT(df_full_PREDICTagetime_qtr) [ , PREDICT_Age_cat_qtr1 := cut(Age_predict_qtr1, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)] 
view(df_full_PREDICTagegroups_qtr1)


df_full_PREDICTagegroups_qtr2 <- setDT(df_full_PREDICTagegroups_qtr1) [ , PREDICT_Age_cat_qtr2 := cut(Age_predict_qtr2, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)] 

view(df_full_PREDICTagegroups_qtr2)


df_full_PREDICTagegroups_qtr3 <- setDT(df_full_PREDICTagegroups_qtr2) [ , PREDICT_Age_cat_qtr3 := cut(Age_predict_qtr3, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)] 

view(df_full_PREDICTagegroups_qtr3)

df_full_PREDICTagegroups_qtr4 <- setDT(df_full_PREDICTagegroups_qtr3) [ , PREDICT_Age_cat_qtr4 := cut(Age_predict_qtr4, 
                                breaks = PEPFARagebreaks19, 
                                right = FALSE, 
                                labels = PEPFARagelabels19)] 

view(df_full_PREDICTagegroups_qtr4)


view(df_full_PREDICTagegroups_qtr4)
glimpse(df_full_PREDICTagegroups_qtr4)
# Rows: 477,120
# Columns: 16
names(df_full_PREDICTagegroups_qtr4)




```





<!-- ############################################################################## -->
<!-- Multiple possible data formats, see archive for 1 & 2 -->
<!-- 1) long with disaggregate (<15 no disaggregate) -->
<!-- 2) long without sex disaggregate -->
<!-- 3) long with all data (male, female and malefemale) - can be disaggregated later -->




###########################################################
######    ALL DATA (format 3) ######
###########################################################


Multiple data formats
1) long with disaggregate
2) long without sex disaggregate
3) long with all data (male, female and malefemale) - can be disaggregated later


## Group by Age Groups and Country and sex
#actual and predicted separately
#merge actual and predicted

```{r load, echo=FALSE}
# 
# #insert the df_long_sexdisagg
# 
# #actual
# df_long_Grouped_Actual_SemiFine_disagg<- df_full_PREDICTagegroups %>%
#   group_by(Country, sex, Time, ACTUAL_Age_cat_SemiFine) %>%
#   summarise(Total_Value_Actual=sum(Value))
# 
# glimpse(df_long_Grouped_Actual_SemiFine_disagg)
# view(df_long_Grouped_Actual_SemiFine_disagg)
# 
# #predict
# df_long_Grouped_Predict_SemiFine_disagg<- df_full_PREDICTagegroups %>%
#   group_by(Country, sex, Time_predict, PREDICT_Age_cat_SemiFine) %>%
#   summarise(Total_Value_Predict=sum(Value))
# 
# glimpse(df_long_Grouped_Predict_SemiFine_disagg)
# view(df_long_Grouped_Predict_SemiFine_disagg)
# 
# 
# #rename predict variables to match actual so they can merge
# df_long_Grouped_Predict_SemiFine_vars_disagg<- df_long_Grouped_Predict_SemiFine_disagg %>%
#   rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine)
# 
# glimpse(df_long_Grouped_Predict_SemiFine_vars_disagg)
# view(df_long_Grouped_Predict_SemiFine_vars_disagg)
# 
# #merge back together
# df_long_Grouped_SemiFine_disagg <- df_long_Grouped_Actual_SemiFine_disagg %>% 
#   left_join(df_long_Grouped_Predict_SemiFine_vars_disagg, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))
# 
# view(df_long_Grouped_SemiFine_disagg)
# # 
# # write_csv(df_long_Grouped_SemiFine_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_grouped_20200914.csv", na="")



```


```{r load, echo=FALSE}


#actual
df_long_Grouped_Actual_SemiFine_disagg<- df_full_PREDICTagegroups_qtr4 %>%
  group_by(Country, sex, Time, ACTUAL_Age_cat_SemiFine) %>%
  summarise(Total_Value_Actual=sum(Value))

glimpse(df_long_Grouped_Actual_SemiFine_disagg)
view(df_long_Grouped_Actual_SemiFine_disagg)




#predict
df_long_Grouped_Predict_SemiFine_disagg1<- df_full_PREDICTagegroups_qtr4 %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_qtr1) %>%
  summarise(Total_Value_Predict_qtr1=sum(Value))%>%
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_qtr1)
view(df_long_Grouped_Predict_SemiFine_disagg1)


df_long_Grouped_Predict_SemiFine_disagg2<- df_full_PREDICTagegroups_qtr4 %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_qtr2) %>%
  summarise(Total_Value_Predict_qtr2=sum(Value)) %>% 
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_qtr2)
  
df_long_Grouped_Predict_SemiFine_disagg3<- df_full_PREDICTagegroups_qtr4 %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_qtr3) %>%
  summarise(Total_Value_Predict_qtr3=sum(Value)) %>% 
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_qtr3)
  
df_long_Grouped_Predict_SemiFine_disagg4<- df_full_PREDICTagegroups_qtr4 %>%
  group_by(Country, sex, Time_predict, PREDICT_Age_cat_qtr4) %>%
  summarise(Total_Value_Predict_qtr4=sum(Value)) %>% 
  rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_qtr4)




#merge back together
df_long_Grouped_SemiFine_disagg <- df_long_Grouped_Actual_SemiFine_disagg %>% 
  left_join(df_long_Grouped_Predict_SemiFine_disagg1, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))%>% 
  left_join(df_long_Grouped_Predict_SemiFine_disagg2, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))%>% 
  left_join(df_long_Grouped_Predict_SemiFine_disagg3, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))%>% 
  left_join(df_long_Grouped_Predict_SemiFine_disagg4, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex"))

view(df_long_Grouped_SemiFine_disagg)
# 
# write_csv(df_long_Grouped_SemiFine_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_grouped_20200914.csv", na="")



```





## look at difference between actual ages - aging in/out actual loss or growth 
#ex actual 2019 / actual 2018

##calc predicted % difference
#predicted 2019/actual 2018

##Compare actual to predicted - expected impact based on aging


```{r load, echo=FALSE}

## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
## This version calculates the perc for all years
df_perc_disagg<- df_long_Grouped_SemiFine_disagg %>%
  janitor::clean_names() %>% 
  arrange(actual_age_cat_semi_fine) %>% #making year values sequential 
  group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs so order to keep lag/lead in these groups
  mutate(total_prev_value_actual_qtr1 = lag(total_value_actual),
  
    perc_of_lastqtr_1 = ifelse(
      !is.na(total_value_predict_qtr1) & !is.na(lag(total_value_actual)),
      total_value_predict_qtr1 / total_prev_value_actual_qtr1 * 100,
      NA),
    
    perc_of_lastqtr_2 = ifelse(
      !is.na(total_value_predict_qtr2 & !is.na(total_value_predict_qtr1)) ,
      total_value_predict_qtr2 / total_value_predict_qtr1 * 100,
      NA),
      
    perc_of_lastqtr_3 = ifelse(
      !is.na(total_value_predict_qtr3& !is.na(total_value_predict_qtr2)) ,
      total_value_predict_qtr3 / total_value_predict_qtr2 * 100,
      NA),
      
    perc_of_lastqtr_4 = ifelse(
      !is.na(total_value_predict_qtr4 & !is.na(total_value_predict_qtr3)) ,
      total_value_predict_qtr4 / total_value_predict_qtr3 * 100,
      NA)
#  ),
 #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100
 ) %>% 
  ungroup()
# %>% 
#   relocate(total_prev_value_actual, .after = total_value_actual)
#   
View(df_perc_disagg)
  


  

```




```{r load, echo=FALSE}

# ## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0
# ## This version calculates the perc for all years
# df_perc_disagg<- df_long_Grouped_SemiFine_disagg %>%
#   janitor::clean_names() %>% 
#   arrange(actual_age_cat_semi_fine) %>% #making year values sequential 
#   group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups
#   mutate(total_prev_value_actual = lag(total_value_actual),
#     
#     perc_diff = ifelse(
#     !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na)
#     total_value_actual / total_prev_value_actual*100, 
#     NA
#     ),
#     
#     perc_of_lastyr = ifelse(
#       !is.na(total_value_predict) & !is.na(lag(total_value_actual)),
#       total_value_predict / total_prev_value_actual * 100,
#       NA
# #  ),
#  #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100
#  )) %>% 
#   ungroup() %>% 
#   relocate(total_prev_value_actual, .after = total_value_actual)
#   
# View(df_perc_disagg)
#   


  

```





#filter ART, time and select relevant variables



```{r load, echo=FALSE}
spectrum_clean_all_disagg<-df_perc_disagg%>%
  filter(time %in% c('2019')) %>%
  select(country, time, sex, actual_age_cat_semi_fine, perc_of_lastyr)%>%
  rename(year=time, age_group=actual_age_cat_semi_fine)%>%
  mutate_if(is.factor, as.character)%>%
  mutate(sex=ifelse(sex==" Female", "Female", sex)) %>%
  mutate(sex=ifelse(sex==" Male", "Male", sex)) %>%
  glimpse()

view(spectrum_clean_all_disagg)
glimpse(spectrum_clean_all_disagg)


write_csv(spectrum_clean_all_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_total_20201123_FY20.csv", na="")

```







#################################   ARCHIVE CODE ##############################

<!-- ## Start with long non-disaggregated because it has all age groups "MaleFemale" -->

<!-- #### SKIP BELOW FOR SEX DISAGGREGATE DATA OR ALL THE WAY TO BOTTOM TOTAL DATASET (MALEFEMALE, MALE, FEMALE) -->


<!-- ########################################################### -->
<!-- ######    NO DISAGGREGATE (format 1) ###### -->
<!-- ########################################################### -->

<!-- ```{r load, echo=FALSE} -->

<!-- #2 -->
<!--   #drop the malefemale==0 and sex -->

<!-- df_long_total <- df_full_PREDICTagegroups %>% -->
<!--     subset(sex="MaleFemale") %>% -->
<!--     select(!sex) %>% -->
<!--     glimpse() -->


<!-- # df_long_total <- df_full_PREDICTagegroups %>% -->
<!-- #     subset(malefemale==1) %>% -->
<!-- #     select(!malefemale) %>% -->
<!-- #     select(!sex) %>% -->
<!-- #     glimpse() -->
<!-- #    -->
<!-- view(df_long_total) -->
<!-- glimpse(df_long_total) -->

<!-- #Rows: 45,360 -->
<!-- #Columns: 9 -->
<!-- #  -->
<!-- # write_csv(df_long_total,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_20200914.csv", na="") -->
<!-- ``` -->



<!-- ## Group by Age Groups and Country  -->
<!-- #actual and predicted separately -->
<!-- #merge actual and predicted -->

<!-- ```{r load, echo=FALSE} -->

<!-- #insert the df_long_total -->

<!-- #actual -->
<!-- df_long_Grouped_Actual_SemiFine<- df_long_total %>% -->
<!--   group_by(Country, Time, ACTUAL_Age_cat_SemiFine) %>% -->
<!--   summarise(Total_Value_Actual=sum(Value)) -->

<!-- glimpse(df_long_Grouped_Actual_SemiFine) -->
<!-- view(df_long_Grouped_Actual_SemiFine) -->

<!-- #predict -->
<!-- df_long_Grouped_Predict_SemiFine<- df_long_total %>% -->
<!--   group_by(Country, Time_predict, PREDICT_Age_cat_SemiFine) %>% -->
<!--   summarise(Total_Value_Predict=sum(Value)) -->

<!-- glimpse(df_long_Grouped_Predict_SemiFine) -->
<!-- view(df_long_Grouped_Predict_SemiFine) -->


<!-- #rename predict variables to match actual so they can merge -->
<!-- df_long_Grouped_Predict_SemiFine_vars<- df_long_Grouped_Predict_SemiFine %>% -->
<!--   rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine) -->

<!-- glimpse(df_long_Grouped_Predict_SemiFine_vars) -->
<!-- view(df_long_Grouped_Predict_SemiFine_vars) -->

<!-- #merge back together -->
<!-- df_long_Grouped_SemiFine <- df_long_Grouped_Actual_SemiFine %>%  -->
<!--   left_join(df_long_Grouped_Predict_SemiFine_vars, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country")) -->

<!-- view(df_long_Grouped_SemiFine) -->





<!-- ``` -->






<!-- ## look at difference between actual ages - aging in/out actual loss or growth  -->
<!-- #ex actual 2019 / actual 2018 -->

<!-- ##calc predicted % difference -->
<!-- #predicted 2019/actual 2018 -->

<!-- ##Compare actual to predicted - expected impact based on aging -->


<!-- ```{r load, echo=FALSE} -->

<!-- ## BK's solution -->
<!-- ## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0 -->
<!-- ## This version calculates the perc for all years -->
<!-- df_perc<- df_long_Grouped_SemiFine %>% -->
<!--   janitor::clean_names() %>%  -->
<!--   arrange(actual_age_cat_semi_fine) %>% #making year values sequential  -->
<!--   group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups -->
<!--   mutate( total_prev_value_actual = lag(total_value_actual), -->

<!--     perc_diff = ifelse( -->
<!--     !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na) -->
<!--     total_value_actual / total_prev_value_actual*100,  -->
<!--     NA -->
<!--     ), -->

<!--     perc_of_lastyr = ifelse( -->
<!--       !is.na(total_value_predict) & !is.na(lag(total_value_actual)), -->
<!--       total_value_predict / total_prev_value_actual * 100, -->
<!--       NA -->
<!-- #  ), -->
<!--  #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100 -->
<!--  )) %>%  -->
<!--   ungroup() %>%  -->
<!--   relocate(total_prev_value_actual, .after = total_value_actual) -->

<!-- View(df_perc) -->





<!-- ``` -->



<!-- #filter ART, time and select relevant variables -->


<!-- ```{r load, echo=FALSE} -->
<!-- spectrum_clean_all<-df_perc%>% -->
<!--   filter(time %in% c( '2019')) %>% -->
<!--   select(country, time, actual_age_cat_semi_fine,   perc_of_lastyr)%>% -->
<!--     rename(year=time, age_group=actual_age_cat_semi_fine)%>% -->
<!--     mutate_if(is.factor, as.character)%>% -->
<!--   glimpse() -->

<!-- view(spectrum_clean_all) -->
<!-- write_csv(spectrum_clean_all,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_all_20200916.csv", na="") -->

<!-- ``` -->










<!-- ########################################################### -->
<!-- ######    SEX DISAGGREGATE (format 2) ###### -->
<!-- ########################################################### -->


<!-- Multiple data formats -->
<!-- 1) long with disaggregate -->
<!-- 2) long without sex disaggregate -->
<!-- 3) long with all data (male, female and malefemale) - can be disaggregated later -->
<!-- ```{r load, echo=FALSE} -->


<!-- #1 -->
<!--   #drop the malefemale==1 -->
<!-- df_long_sexdisagg <- df_full_PREDICTagegroups %>% -->
<!--     subset(sex!=2) %>% -->
<!--     view() -->

<!-- view(df_long_sexdisagg) -->
<!-- glimpse(df_long_sexdisagg) -->
<!-- #Rows: 73,920 -->
<!-- #Columns: 10 -->

<!-- # write_csv(df_long_sexdisagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/Spectrum_long_sexdisagg_20200914.csv", na="") -->

<!-- ``` -->





<!-- ## Group by Age Groups and Country and sex -->
<!-- #actual and predicted separately -->
<!-- #merge actual and predicted -->

<!-- ```{r load, echo=FALSE} -->

<!-- #insert the df_long_sexdisagg -->

<!-- #actual -->
<!-- df_long_Grouped_Actual_SemiFine_disagg<- df_long_sexdisagg %>% -->
<!--   group_by(Country, sex, Time, ACTUAL_Age_cat_SemiFine) %>% -->
<!--   summarise(Total_Value_Actual=sum(Value)) -->

<!-- glimpse(df_long_Grouped_Actual_SemiFine_disagg) -->
<!-- view(df_long_Grouped_Actual_SemiFine_disagg) -->

<!-- #predict -->
<!-- df_long_Grouped_Predict_SemiFine_disagg<- df_long_sexdisagg %>% -->
<!--   group_by(Country, sex, Time_predict, PREDICT_Age_cat_SemiFine) %>% -->
<!--   summarise(Total_Value_Predict=sum(Value)) -->

<!-- glimpse(df_long_Grouped_Predict_SemiFine_disagg) -->
<!-- view(df_long_Grouped_Predict_SemiFine_disagg) -->


<!-- #rename predict variables to match actual so they can merge -->
<!-- df_long_Grouped_Predict_SemiFine_vars_disagg<- df_long_Grouped_Predict_SemiFine_disagg %>% -->
<!--   rename(Time=Time_predict, ACTUAL_Age_cat_SemiFine=PREDICT_Age_cat_SemiFine) -->

<!-- glimpse(df_long_Grouped_Predict_SemiFine_vars_disagg) -->
<!-- view(df_long_Grouped_Predict_SemiFine_vars_disagg) -->

<!-- #merge back together -->
<!-- df_long_Grouped_SemiFine_disagg <- df_long_Grouped_Actual_SemiFine_disagg %>%  -->
<!--   left_join(df_long_Grouped_Predict_SemiFine_vars_disagg, by=c("Time"="Time", "ACTUAL_Age_cat_SemiFine"="ACTUAL_Age_cat_SemiFine","Country"="Country", "sex"="sex")) -->

<!-- view(df_long_Grouped_SemiFine_disagg) -->





<!-- ``` -->






<!-- ## look at difference between actual ages - aging in/out actual loss or growth  -->
<!-- #ex actual 2019 / actual 2018 -->

<!-- ##calc predicted % difference -->
<!-- #predicted 2019/actual 2018 -->

<!-- ##Compare actual to predicted - expected impact based on aging -->


<!-- ```{r load, echo=FALSE} -->

<!-- ## BK's solution -->
<!-- ## Note: Make sure to install `janitor` and `dplyr` version >= 1.0.0 -->
<!-- ## This version calculates the perc for all years -->
<!-- df_perc_disagg<- df_long_Grouped_SemiFine_disagg %>% -->
<!--   janitor::clean_names() %>%  -->
<!--   arrange( actual_age_cat_semi_fine) %>% #making year values sequential  -->
<!--   group_by(country, actual_age_cat_semi_fine) %>% #dont want to confuse the countries or hiv/art obs - keep lag/lead in these groups -->
<!--   mutate( total_prev_value_actual = lag(total_value_actual), -->

<!--     perc_diff = ifelse( -->
<!--     !is.na(total_value_actual) & !is.na(total_prev_value_actual), #ignore the values that are empty (na) -->
<!--     total_value_actual / total_prev_value_actual*100,  -->
<!--     NA -->
<!--     ), -->

<!--     perc_of_lastyr = ifelse( -->
<!--       !is.na(total_value_predict) & !is.na(lag(total_value_actual)), -->
<!--       total_value_predict / total_prev_value_actual * 100, -->
<!--       NA -->
<!-- #  ), -->
<!--  #       perc_diff_Compare2= (perc_diff/perc_of_lastyr)*100 -->
<!--  )) %>%  -->
<!--   ungroup() %>%  -->
<!--   relocate(total_prev_value_actual, .after = total_value_actual) -->

<!-- View(df_perc_disagg) -->





<!-- ``` -->



<!-- #filter ART, time and select relevant variables -->


<!-- ```{r load, echo=FALSE} -->
<!-- spectrum_clean_disagg<-df_perc_disagg%>% -->
<!--   filter(time %in% c('2019')) %>% -->
<!--   select(country, time, sex, actual_age_cat_semi_fine, perc_of_lastyr)%>% -->
<!--   rename(year=time, age_group=actual_age_cat_semi_fine)%>% -->
<!--   mutate_if(is.factor, as.character)%>% -->
<!--   mutate(sex=ifelse(sex==" Female", "Female", "Male")) %>% -->
<!--   glimpse() -->

<!-- view(spectrum_clean_disagg) -->
<!-- glimpse(spectrum_clean_disagg) -->

<!-- write_csv(spectrum_clean_disagg,"C:/Users/jesse/OneDrive/Documents/Work Material/USAID/Treatment/Data/Output/spectrum_clean_sexdisagg_20200916.csv", na = "") -->

<!-- ``` -->













### Aditional Archives


<!-- # Figure out predicted ages  -->
<!-- #ex, 2017 ages if 2016 aged w/o other loss/gains -->


<!-- ```{r load, echo=FALSE} -->



<!-- PREDICTagebreaks <- c(0,4,9,14,19,24,29,34,39,44,49,85) -->
<!-- PREDICTagelabels <- c("0-3","4-8","9-13","14-18","19-23","24-28","29-33", -->
<!--                "34-38","39-43","44-48","49+") -->

<!-- PREDICTagebreaks18 <- c(0,9,14,19,24,29,34,39,44,49,85) -->
<!-- PREDICTagelabels18 <- c("0-8","9-13","14-18","19-23","24-28","29-33", -->
<!--                "34-38","39-43","44-48","49+") -->

<!-- df_wide_PREDICTagegroupsFINE <- setDT(df_combined_PEPFARagegroups) [ , PREDICT_Age_cat_Fine := cut(Age_num,  -->
<!--                                 breaks = PREDICTagebreaks,  -->
<!--                                 right = FALSE,  -->
<!--                                 labels = PREDICTagelabels)] -->
<!-- view(df_wide_PREDICTagegroupsFINE) -->
<!-- glimpse(df_wide_PREDICTagegroupsFINE) -->

<!-- df_wide_PREDICTagegroupsALL <- setDT(df_wide_PREDICTagegroupsFINE) [ , PREDICT_Age_cat_SemiFine := cut(Age_num,  -->
<!--                                 breaks = PREDICTagebreaks18,  -->
<!--                                 right = FALSE,  -->
<!--                                 labels = PREDICTagelabels18)] -->
<!-- view(df_wide_PREDICTagegroupsALL) -->
<!-- glimpse(df_wide_PREDICTagegroupsALL) -->

<!-- ``` -->



<!-- ## Figure out predicted ages  -->
<!-- #ex, 2017 ages if 2016 aged w/o other loss/gains -->


<!-- ```{r load, echo=FALSE} -->


<!-- for each year and actual age cat -->
<!--   do value of previous year and  -->


<!-- PRED_Value_2019(ACTUAL_Age_cat_Fine) = ACTUAL_Value_2018 (PRED_age_bands) -->


<!-- #loop over age bands -->

<!-- #long -->

<!-- for(i in Value){ -->
<!--   Pred_Value <- when Time=Time-1 & PRED_Age_cat_SemiFine=0 -->



<!--  # wide??? -->

<!--   sum(ACTUAL_Value_2018) if PRED_age_bands =  -->
<!-- } -->

<!-- PRED_Value_2019_1-9  <- ACTUAL_Value_2018_0-8 -->

<!-- PRED_Value_2019_1-9  <- ACTUAL_Value_2018_0-8 -->


<!-- ``` -->



